\Large \bf Vigor: Toward Carrier-Grade Software Network Functions

### 2016-11-21 doc/paper/paper.tex
Luis's feedback: Expanded captions for listings; short explanation of havocing; some line number placeholders.

  while(1) {
    if (queue_len < CAP) {
      int end = (begin + queue_len)%CAP;
      if (receive_packet(&packets[end]) &&
          packets[end].port != 9)
        queue_len += 1;
    }
    if (0 < queue_len) {
      send_packet(&packets[begin]);
      begin = (begin + 1)%CAP;
    }
  }
  return 0;
}
\end{lstlisting}
  \caption{Original implementation of the Drop protocol}
  \label{lst:orig}
\end{figure}

@@

{\bf Step 2: Use library components. } A developer replaces every array with a
component from Vigor library that performs the same function. For our example,
the library provides a packet ring with a simple interface: \code{bool
  ring_empty(ring*)} checks whether there is a packet in the ring, \code{bool
  ring_full(ring*)} checks whether the ring has run out of space, \code{void
  ring_push_back(ring*, packet* src)} adds packet \code{src} to the ring, and
\code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
and places it in \code{dst}. Fig.\ref{lst:vigor} shows the resulting NF code.
\code{VIGOR_CHECK} adds additional property that we would like to verify ---
that we do not forward packets with port 9.

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);
  if (!r) return 1;
  while(LOOP(1))
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);
      VIGOR_CHECK(p.port != 9);
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Vigor implementation of Drop protocol.}
  \label{lst:vigor}
\end{figure}

@@

static bool
packet_constraints(struct packet* p) {
  return p->port != 9;
}
\end{lstlisting}
  \caption{Custom definitions capturing the desired property.}
  \label{lst:customizations}
\end{figure}

Currently, we have to duplicate the packet condition in two languages (formal
logic, and C) due to the limitations of our prototype.

{\bf Step 4: Run Vigor checker.} Vigor will find the appropriate symbolic model,
check that the application does not crash in any execution scenario. It will
then make sure that the chosen model is faithful to the formally proven
implementation of the component.

@@

Vigor transforms infinite loops using loop
invariants~\cite[\S~2.1]{cormen2009introduction}. We added to KLEE support for
automatic havoc-ing~\cite{barnett2005boogie} (replace all variables changed in
the loop with unconstrained symbols). The step 1 enables Vigor to reduce each
problematic loop to a single generalized iteration that allows reasoning in one
pass.

@@

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p)
/*@ requires ringp(r, ?user_property,
                   ?lst, ?cap) &*&
             lst != nil &*& packetp(p, _); @*/
/*@ ensures ringp(r, user_property,
                  tail(lst), cap) &*&
            packetp(p, head(lst)) &*&
            true == user_property(head(lst)); @*/
{
  //@ extract_first(r);
  struct packet* src_pkt = r->array + r->begin;
  p->port = src_pkt->port;
  r->len = r->len - 1;
  r->begin = r->begin + 1;
  if (r->cap <= r->begin) {
    r->begin = 0;
    //@ stitch_with_empty_overflow(r);
  } else {
    //@ stitch_with_empty(r);
  }
}
\end{lstlisting}
  \caption{A formal contract and verified implementation of the \code{ring_pop_front} function.}
  \label{lst:contract}
\end{figure}

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  ASSUME(packet_constraints(p));
}
\end{lstlisting}
  \caption{A symbolic model for \code{ring_pop_front}.}
  \label{lst:symbolic-model}
\end{figure}

@@

You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
It starts with \code{loop_invariant_produce} that marks the beginning of a loop
iteration. It continues then with calls to \code{ring_full} \code{ring_empty},
noting (with \code{assume}) that both return \code{false}. The seqeunce ends
with a call to \code{ring_pop_front}. The last line detects underapproximation
of the model output.

@@

/*@ assert(arg2.port != 9); @*/
\end{lstlisting}
  \caption{A call sequence, produced by SEE.}
  \label{lst:call-sequence}
\end{figure}

@@

Consider potential model mismatches. Suppose we had an overapproximate model ---
Fig.~\ref{lst:overapproximate}. Then the \code{VIGOR_CHECK}
(Fig.\ref{lst:vigor}) in the application code would fail.
\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  // No constraints on the packet port here.
}
\end{lstlisting}
  \caption{An overapproximate ring model fragment.}
  \label{lst:overapproximate}
\end{figure}

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  p->port = 0;
}
\end{lstlisting}
  \caption{An underapproximate ring model fragment.}
  \label{lst:underapproximate}
\end{figure}

>>>>

  while(1) {
    if (queue_len < CAP) {
      int end = (begin + queue_len)%CAP;
      if (receive_packet(&packets[end]) &&
          packets[end].port != 9)
        queue_len += 1;
    }
    if (0 < queue_len) {
      send_packet(&packets[begin]);
      begin = (begin + 1)%CAP;
    }
  }
  return 0;
}
\end{lstlisting}
  \caption{Original implementation of the Drop protocol. Contains an infinite
    loop and a latent ring implementation}
  \label{lst:orig}
\end{figure}

@@

{\bf Step 2: Use library components. } A developer replaces every array with a
component from Vigor library that performs the same function. For our example,
the library provides a packet ring with a simple interface: \code{bool
  ring_empty(ring*)} checks whether there is a packet in the ring, \code{bool
  ring_full(ring*)} checks whether the ring has run out of space, \code{void
  ring_push_back(ring*, packet* src)} adds packet \code{src} to the ring, and
\code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
and places it in \code{dst}. Fig.\ref{lst:vigor} shows the resulting NF code.
\code{VIGOR_CHECK} (line NN) adds additional property that we would like to
verify --- that we do not forward packets with port 9.

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);
  if (!r) return 1;
  while(LOOP(1))
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);
      VIGOR_CHECK(p.port != 9);
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Vigor implementation of Drop protocol. \code{LOOP} labels the
    infinite loop head and we explicitly use a ring from the Vigor library.
    \code{VIGOR_CHECK} at the line NN adds a semantic property to be verified.}
  \label{lst:vigor}
\end{figure}

@@

static bool
packet_constraints(struct packet* p) {
  return p->port != 9;
}
\end{lstlisting}
  \caption{Custom definitions capturing the desired property. They capture the
    essential property of every packet in the ring.}
  \label{lst:customizations}
\end{figure}

Currently, we have to duplicate the packet condition in two languages (formal
logic -lines NN, and C - NN) due to the limitations of our prototype.

{\bf Step 4: Run Vigor checker.} Vigor will find the appropriate symbolic model
and check that the application does not crash in any execution scenario. It will
then make sure that the chosen model is faithful to the formally proven
implementation of the component.

@@

Vigor transforms infinite loops using loop
invariants~\cite[\S~2.1]{cormen2009introduction}. We added to KLEE support for
automatic havoc-ing~\cite{barnett2005boogie}. Havoc-ing simply replaces all the
variables that may change in the loop by unconstrained symbols at the beginning
of the iteration. In our example that would be the variables \code{begin} and
\code{queue_len} and the contents of \code{packets[0..CAP]}, but not the value
of the \code{packets} pointer because it never changes. The step 1 enables
Vigor to reduce each problematic loop to a single generalized iteration that
allows reasoning in one pass.

@@

In this example the symbolic indexing comes from the ring array. Once we
replaced the infinite loop with a single iteration we have to assume (remember
havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may point
anywhere in the array.

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p)
/*@ requires ringp(r, ?user_property,
                   ?lst, ?cap) &*&
             lst != nil &*& packetp(p, _); @*/
/*@ ensures ringp(r, user_property,
                  tail(lst), cap) &*&
            packetp(p, head(lst)) &*&
            true == user_property(head(lst)); @*/
{
  //@ extract_first(r);
  struct packet* src_pkt = r->array + r->begin;
  p->port = src_pkt->port;
  r->len = r->len - 1;
  r->begin = r->begin + 1;
  if (r->cap <= r->begin) {
    r->begin = 0;
    //@ stitch_with_empty_overflow(r);
  } else {
    //@ stitch_with_empty(r);
  }
}
\end{lstlisting}
  \caption{A formal contract and verified implementation of the
    \code{ring_pop_front} function. The comments starting with \code{@} symbol
    are formal annotations that hold the function contract (\code{requires} -
    precondition, \code{ensures} - postcondition) and lemmas that help the proof
    checker to verify the state transformations in the code.}
  \label{lst:contract}
\end{figure}

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  ASSUME(packet_constraints(p));
}
\end{lstlisting}
  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically the
    a packet and adds a constraint to it.}
  \label{lst:symbolic-model}
\end{figure}

@@

You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
It starts with \code{loop_invariant_produce} that marks the beginning of a loop
iteration. It continues then with calls to \code{ring_full} and \code{ring_empty},
noting (with \code{assume}) that both return \code{false}. The seqeunce ends
with a call to \code{ring_pop_front}. The last line detects underapproximation
of the model output.

@@

/*@ assert(arg2.port != 9); @*/
\end{lstlisting}
  \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
    constrain the execution space.}
  \label{lst:call-sequence}
\end{figure}

@@

Consider potential model mismatches. Suppose we had an overapproximate model ---
Fig.~\ref{lst:overapproximate}. Then the \code{VIGOR_CHECK}
(Fig.\ref{lst:vigor}) in the application code would fail.
\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  // No constraints on the packet port here.
}
\end{lstlisting}
  \caption{An overapproximate ring model fragment. It does not add the
    constraint necessary for the code to verify.}
  \label{lst:overapproximate}
\end{figure}

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  p->port = 0;
}
\end{lstlisting}
  \caption{An underapproximate ring model fragment. Instead of returning a
    constrained symbolic packet, it simply returns some particular packet, which
    clearly can not capture what the real ring would return on its place.}
  \label{lst:underapproximate}
\end{figure}
